Using reference models in feature selection problems \vspace{.1in}

### 2019-07-22 paper/ref_approach.tex
minor edits

\cite{paper:bodyfat} report results using steplm with a significance level of 0.157 with AIC selection \citep{akaike1974new}, fixing abdomen and height to be always included in the model. Here, we implement steplm via AIC without fixing any variable. For selection via projpred, the reference model includes all the covariates, we regularise model complexity with a regularised horseshoe prior on the covariate coefficients \citep{paper:rhs}, submodels are explored using forward search, and the predictive utility is the expected log-predictive density (elpd) estimated using PSIS-LOO \citep{paper:psis_loo}. The experiments are implemented in \texttt{R} \citep{Rcore2018}. The steplm approach is carried out combining the \texttt{step} and \texttt{lm} functions, whereas the selection via projpred is done with the \texttt{projpred} package. In order to automatise the latter, we select the submodel as the one with the smallest size which has an elpd score higher (i.e., better) than the reference model score with probability of at least 5\%. The inclusion frequencies of each feature in the final model after 100 bootstrap samples are depicted in Figure \ref{fig:inclusion_frequencies}. The projpred approach has only two variables, 'abdomen' and 'weight', with inclusion frequencies above 50\% ('abdomen' is the only included always), the third most included is `wrist' at 44\%, and the fourth is `height' at 35\%, while steplm has seven features above 50\%. Such a lower stability of steplm can be observed also in the bootstrap model selection frequencies reported in Table \ref{tab:model_frequencies}. For example, the first five selected models have a cumulative frequency of 76\% with projpred, but only of 14\% with steplm. In addition, we note that the size of the selected model is much smaller when using projpred.

@@

We consider the reference model approach as a family of methods, thus providing a general answer to the question of computational burden is not possible because not unique. It is possible to identify a common and inevitable overhead, though, which is the cost of fitting the reference model. In the simple ''bare-bones'' reference model approach we have described above, it is actually the only added cost. In general, a full Bayesian model is advisable in order to have the best predictive performance. However, including all the available features can often be quite computationally demanding. In such a case, a first possible speed-up could be using screening or dimensionality reduction techniques, as for example the supervised principal components \citep{bair2006prediction,piironen2018}. However, in some cases, this does not always result in an appreciable speed-up. For example, when there are several predictive but mutually uncorrelated features, all of them  have to be included in the reference model anyway. Other alternatives could be fixing some parameters to their maximum marginal likelihood value \cite[typically referred as empirical Bayes;][]{casella1985introduction} or use a completely frequentist approach, if predictions are good enough.

@@

Within the set of complete selection procedures, we consider three 
Bayesian methods which are described in the following.
\cite{johnstone2004needles} use a spike-and-slab type prior given by a mixture of a delta distribution in zero and a heavy-tailed distribution \cite[see also][]{paper:spike_slab_mitchell}. When the latter is a Laplace distribution, they show an interesting thresholding property of the median estimator and, based on this observation, present a complete selection procedure using their prior. We refer to this method as EB.med. Further, we use a method to control for the local false discovery rate (loc.fdr) as described in \cite{paper:efron, efron2012large}. Further, we device a simple selection procedure based on posterior credible intervals using the regularised horseshoe prior in which we select covariates whose 90\% credible intervals excludes zero (ci.90). All of these methods provide a complete selection procedure and we compare their performance with and without using a reference model on top of them. As criteria of selection performance, we evaluate the average false discovery rate (i.e., the ratio of the number of non-relevant selected features over the number of selected features) and the average sensitivity (i.e., the ratio of the number of relevant selected features over the total number of relevant features). We also provide a comparison of the stability of the selection by means of a stability measure proposed by \cite{paper:stability}. Table \ref{tab:comparison} summarises the methods used in out numerical experiments.

@@

We conclude our experiments using the real world 'body fat' dataset one more time, which was already described in the comparison between the projection predictive approach and the stepwise regression in Section \ref{projection}. We carry out two different analyses: One using the normal means problem framework and another one using selection via stepwise backward regression. In the former, we add noisy uncorrelated covariates to the original data up to a total of 1000 features. These covariates are normally distributed and scaled as the original covariates. We compute correlations between each variable and the target variable, that is the amount of fat, and transform them by Fisher-Z-transformation. The original assumption in order for \eqref{eq:fisher_transformation} to hold is that the variables are jointly normally distributed. In our experience the normal approximation in \eqref{eq:fisher_transformation} is still reasonable, but after rescaling by $\sqrt{n-3}$ we do not fix the variance to be one, and instead estimate it from the data. The methods we compare are those used in Section \ref{complete-selection}: the control of the local false discovery rate (loc.fdr), the empirical Bayes median (EB.med) and the selection by posterior credible intervals at level 90\% (ci.90). 
In order to vary the difficulty of the selection, we bootstrap subsamples of different sizes, going from $n=50$ up to $n=251$ (i.e., the full size of the data). For each condition, results are averaged over 100 bootstrap samples of the respective size.

>>>>

\cite{paper:bodyfat} report results using steplm with a significance level of 0.157 with AIC selection \citep{akaike1974new}, fixing abdomen and height to be always included in the model. Here, we implement steplm via AIC without fixing any variable. For the selection via projpred, the reference model includes all the covariates, we regularise model complexity with a regularised horseshoe prior on the covariate coefficients \citep{paper:rhs}, submodels are explored using forward search, and the predictive utility is the expected log-predictive density (elpd) estimated using PSIS-LOO \citep{paper:psis_loo}. The experiments are implemented in \texttt{R} \citep{Rcore2018}. The steplm approach is carried out combining the \texttt{step} and \texttt{lm} functions, whereas the selection via projpred is done with the \texttt{projpred} package. In order to automatise the latter, we select the submodel as the one with the smallest size which has an elpd score higher (i.e., better) than the reference model score with probability of at least 5\%. The inclusion frequencies of each feature in the final model after 100 bootstrap samples are depicted in Figure \ref{fig:inclusion_frequencies}. The projpred approach has only two variables, `abdomen' and `weight', with inclusion frequencies above 50\% (`abdomen' is the only included always), the third most included is `wrist' at 44\%, and the fourth is `height' at 35\%, while steplm has seven features above 50\%. Such a lower stability of steplm can be observed also in the bootstrap model selection frequencies reported in Table \ref{tab:model_frequencies}. For example, the first five selected models have a cumulative frequency of 76\% with projpred, but only of 14\% with steplm. In addition, we note that the size of the selected model is much smaller when using projpred.

@@

We consider the reference model approach as a family of methods, thus providing a general answer to the question of computational burden is not possible because not unique. It is possible to identify a common and inevitable overhead, though, which is the cost of fitting the reference model. In the simple ``bare-bones'' reference model approach we have described above, it is actually the only added cost. In general, a full Bayesian model is advisable in order to have the best predictive performance. However, including all the available features can often be quite computationally demanding. In such a case, a first possible speed-up could be using screening or dimensionality reduction techniques, as for example the supervised principal components \citep{bair2006prediction,piironen2018}. However, in some cases, this does not always result in an appreciable speed-up. For example, when there are several predictive but mutually uncorrelated features, all of them  have to be included in the reference model anyway. Other alternatives could be fixing some parameters to their maximum marginal likelihood value \cite[typically referred as empirical Bayes;][]{casella1985introduction} or use a completely frequentist approach, if predictions are good enough.

@@

Within the set of complete selection procedures, we consider three 
Bayesian methods which are described in the following.
\cite{johnstone2004needles} use a spike-and-slab type prior given by a mixture of a delta distribution in zero and a heavy-tailed distribution \cite[see also][]{paper:spike_slab_mitchell}. When the latter is a Laplace distribution, they show an interesting thresholding property of the median estimator and, based on this observation, present a complete selection procedure using their prior. We refer to this method as EB.med. Further, we use a method to control for the local false discovery rate (loc.fdr) as described in \cite{paper:efron, efron2012large}. Finally, we device a simple selection procedure based on posterior credible intervals using the regularised horseshoe prior in which we select covariates whose 90\% credible intervals excludes zero (ci.90). All of these methods provide a complete selection procedure and we compare their performance with and without using a reference model on top of them. As criteria of selection performance, we evaluate the average false discovery rate (i.e., the ratio of the number of non-relevant selected features over the number of selected features) and the average sensitivity (i.e., the ratio of the number of relevant selected features over the total number of relevant features). We also provide a comparison of the stability of the selection by means of a stability measure proposed by \cite{paper:stability}. Table \ref{tab:comparison} summarises the methods used in out numerical experiments.

@@

We conclude our experiments using the real world `body fat' dataset one more time, which was already described in the comparison between the projection predictive approach and the stepwise regression in Section \ref{projection}. We carry out two different analyses: one using the normal means problem framework and another one using selection via stepwise backward regression. In the former, we add noisy uncorrelated covariates to the original data up to a total of 1000 features. These covariates are normally distributed and scaled as the original covariates. We compute correlations between each variable and the target variable, that is the amount of fat, and transform them by Fisher-Z-transformation. The original assumption in order for \eqref{eq:fisher_transformation} to hold is that the variables are jointly normally distributed. In our experience the normal approximation in \eqref{eq:fisher_transformation} is still reasonable, but after rescaling by $\sqrt{n-3}$ we do not fix the variance to be one, and instead estimate it from the data. The methods we compare are those used in Section \ref{complete-selection}: the control of the local false discovery rate (loc.fdr), the empirical Bayes median (EB.med) and the selection by posterior credible intervals at level 90\% (ci.90). 
In order to vary the difficulty of the selection, we bootstrap subsamples of different sizes, going from $n=50$ up to $n=251$ (i.e., the full size of the data). For each condition, results are averaged over 100 bootstrap samples of the respective size.
